---
title: "Lesson 1 - Data visualization"
output: html_document
date: "2023-05-04"
---

# Lesson 1 - Data visualization {.tabset}

## 1 Importing and viewing the data {.tabset}

*1.1 install tidyverse package*

```{r packages, warning=FALSE,results="hide"}
# tidyverse (8 packages) : import, transform, tidy, and visualize 
library(tidyverse)

```

*1.2 install and load palmerpenguins data frame*

```{r}
library(palmerpenguins)

penguins
```

*1.3 overview of the data : glimpse and skim*

```{r}
glimpse(penguins)

library(skimr)

skim(penguins)
```

*1.4 view the data*

```{r}
view(penguins)
```

## 2 Data visualization with ggplot2 {.tabset}

```{r}
#our goal
data_viz_goal<-ggplot(penguins, 
       aes(x = flipper_length_mm, y = body_mass_g)) +
  geom_point(aes(color = species, shape = species)) +
  geom_smooth() +
  labs(
    title = "Body mass and flipper length",
    subtitle = "Dimensions for Adelie, Chinstrap, and Gentoo Penguins",
    x = "Flipper length (mm)", 
    y = "Body mass (mm)",
    color = "Species", 
    shape = "Species")
```

*2.1 Creating at ggplot*

```{r}
ggplot(data = penguins,
       mapping = aes(x=flipper_length_mm, y=body_mass_g))+
  geom_point()

#we will address the warning message next lesson :)
penguins %>% 
  select(species, flipper_length_mm, body_mass_g) %>% 
  filter(is.na(body_mass_g) | is.na(flipper_length_mm))
```

*2.2 Adding aesthetics and layers*

*2.2.1 Adding color*

```{r}

ggplot(data = penguins,
       mapping = aes(x = flipper_length_mm, y = body_mass_g,
                     color = species)) +
  geom_point()
```

*2.2.2 Adding a smooth curve*

```{r}
ggplot(data = penguins,
       mapping = aes(x = flipper_length_mm, y = body_mass_g,
                     color = species)) +  #color at global level
  geom_point() +
  geom_smooth()


ggplot(data = penguins,
       mapping = aes(x = flipper_length_mm, y = body_mass_g)) +  
  geom_point(mapping = aes(color = species)) + #color at local level
  geom_smooth()
```

*2.2.3 Adding shapes*

```{r}
ggplot(data = penguins,
       mapping = aes(x = flipper_length_mm, y = body_mass_g)) +
  geom_point(mapping = aes(color = species, shape = species)) +
  geom_smooth()

```

*2.2.4 Adding titles*

```{r}
ggplot(penguins, 
       aes(x = flipper_length_mm, y = body_mass_g)) +
  geom_point(aes(color = species, shape = species)) +
  geom_smooth() +
  labs(
    title = "Body mass and flipper length",
    subtitle = "Dimensions for Adelie, Chinstrap, and Gentoo Penguins",
    x = "Flipper length (mm)", 
    y = "Body mass (mm)",
    color = "Species", 
    shape = "Species")
```

*2.3 Visualizing distributions*

*2.3.1 Categorical variables*

```{r}
ggplot(penguins,
       aes(x=species))+
  geom_bar()

  
ggplot(penguins,
       aes(x=fct_infreq(species)))+
  geom_bar()

```

*2.3.2 Numeric variables*

```{r}
#density plot
ggplot(penguins, aes(x = body_mass_g)) +
  geom_density()

# histogram
ggplot(penguins, aes(x = body_mass_g)) +
  geom_histogram(binwidth = 200)
```

*2.4 Visualizing relationships*

*2.4.1 Numerical and categorical variables*

```{r}
#boxplot

ggplot(penguins, aes(x = species, y = body_mass_g)) +
  geom_boxplot()

#density plot
ggplot(penguins, aes(x = body_mass_g, color = species, fill = species)) +
  geom_density(alpha = 0.5)
```

*2.4.2 Two categorical variables*

```{r}
ggplot(penguins, aes(x = island, fill = species)) +
  geom_bar()

ggplot(penguins, aes(x = island, fill = species)) +
  geom_bar(position = "fill")

```

*2.4.3 Two numerical variables*

```{r}
ggplot(penguins, aes(x = flipper_length_mm, y = body_mass_g)) +
  geom_point()

```

*2.4.4 Three of more variables*

```{r}
ggplot(penguins, aes(x = flipper_length_mm, y = body_mass_g)) +
  geom_point(aes(color = species, shape = island))

ggplot(penguins, aes(x = flipper_length_mm, y = body_mass_g)) +
  geom_point(aes(color = species, shape = species))+
  facet_wrap(~island)

```
